Using reference models in feature selection \vspace{.1in}

### 2019-12-07 paper/ref_approach.tex
next round of proofreading

    \author{
    Federico Pavone, 
  Juho Piironen,
  Paul-Christian B\"{u}rkner
  and Aki Vehtari \footnote{Department of Computer Science, Aalto University, Finland.}
  }

@@

\begin{document}
\maketitle
\begin{abstract}
  Feature selection, or more generally, model reduction is an important aspect of the statistical workflow aiming to provide insights from data. In this paper we discuss and demonstrate the benefits of using a reference model in feature selection. A reference model acts as a noise-filter on the target variable by modeling its data generating mechanism. As a result, using the reference model predictions in the model selection procedure reduces the variability and improves stability leading to improved model selection performance. Assuming that a Bayesian reference model describes the true distribution of future data well, the theoretically preferred usage of the reference model is to project its predictive distribution to a reduced model leading to projection predictive variable selection approach. Alternatively, reference models may also be used in an ad-hoc manner in combination with common feature selection methods.  In several numerical experiments, we demonstrate superior performance of projection predictive approach and show how use of a reference model translates into better and more stable feature selection independently of the specific selection method. 
% These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.
\end{abstract}

@@

In statistical applications, one of the main steps in the modelling
workflow is covariate or feature selection, which is a special case of
model reduction. Feature (variable) selection may have multiple goals.  First,
if the covariates themselves are of interest, we can use variable
selection to infer which variables contain predictive information
about the target variable.  Second, as simpler models come with the
advantages of reduced measurement costs and improved interpretability,
we may be interested in finding the minimal subset of covariates which
still provides good predictive performance (or good balance between
simplicity and predictive performance).  When the predictive
capability is guiding the selection, the true data generation
mechanism of future data can be approximated either by using the
observed data directly or alternatively by using predictions from a
\emph{reference model} \citep{vehtari2012survey}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all covariates.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects our beliefs about
the future data in the best possible way and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian,gabry2019visualization}.  The reference model approach
has been used in Bayesian statistics at least since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see the papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see a paper by \citet{paper:projpred}.

Reference models have been also used in non-Bayesian contexts, where
\cite{harrell2015regression} describes them as full models that can be
thought of as a ``gold standard'' (for a given application).  For
example, \cite{faraggi2001understanding} deal with the necessity of
identifying interpretable risk groups in the context of survival data
using neural networks, which typically perform very well in terms of
prediction, but whose covariates are difficult to be understood in
terms of relevance.  \cite{paul2008preconditioning}, using the term
preconditioning, explore approximating models fitting Lasso or
stepwise regression against consistent estimates $\hat{y}$ of a
reference model instead of the observed responses $y$.
Whatever the terminology or applied statistical framework, reference
models offer a powerful approach to improving feature selection as we
will demonstrate in the present paper.

\section{Reference models in feature selection}
\label{intro-example}

@@

\subsection{Bodyfat example: Part 1}
\label{intro-example}

To motivate the further discussion and experiments, we start by simple
variable selection example using body fat data by \citet{johnson1996fitting}.
We compare projection predictive approach (\emph{projpred},
\citet{paper:projpred}) which uses a reference model, and classic
stepwise backward regression (steplm).
% \footnote{The analysis is based on \texttt{R}-notebook by Vehtari available at
%   \url{https://avehtari.github.io/modelselection/bodyfat.html}.}.
The experiments are implemented in \texttt{R} \citep{Rcore2018}.

The target variable of interest is the amount of body fat, which is
measured by immersing a person in water tank which is a complex and
expensive procedure. Additionally, we have information about 13
covariates which are anthropometric measurements (e.g., height, weight
and circumference of different body parts). The
covariates are highly correlated which causes additional
challenge in the feature selection. In total, we have 251
observations. The goal is to find the model which is able to predict
the amount of body fat well while requiring least amount of
measurements for a new person.

\cite{paper:bodyfat} report results using \emph{steplm} with a significance
level of 0.157 with AIC selection \citep{akaike1974new}, fixing
abdomen and height to be always included in the model. For better
comparison we don't fix any of the variables.  The steplm approach is
carried out combining the \texttt{step} and \texttt{lm} functions.

For the selection via projpred, the Bayesian reference model includes
all the covariates using a regularised horseshoe prior~\citep{paper:rhs} on the
covariate coefficients. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using approximate leave-one-out cross-validation via Pareto-smoothed importance-sampling
\citep[PSIS-LOO-CV; ][]{paper:psis_loo}.  We select the smallest  submodel with an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach, and papers by \citet{paper:model_selection} and
\citet{paper:projpred} for more details. The complete projpred approach is implemented in the \texttt{projpred} R package.

@@

We repeat the experiment with a modified data set by adding 84
unrelated noisy features. The last two rows of Table
\ref{tab:model_performances} show the cross-validated RMSE, the size
of the selected model and the number of selected noisy features using
projpred or steplm.  The projection predictive approach has similar
predictive performance and the same number of selected features as
with the original data, whereas the stepwise regression has worse
predictive performance and the number of selected features is much
higher and include a large number of irrelevant features.

@@

Both projpred and steplm compare a large number of models using either
forward or backward search, which can lead to selection induced
overfitting, but even with 100 covariates, projpred is able to select a
submodel with similar performance as the full model. In this example, the two
 compared methods also differ in other aspects than the usage of a reference 
 model, such as that projpred
uses Bayesian inference and steplm uses maximum likelihood estimation. 
However, as we show in the present paper, one of the primary difference 
with respect to quality of the variable selection is indeed whether or 
not a reference model approach is applied.

@@

If one of the models to be compared is the full model which can be
used as a reference model, there is no additional cost of using a
reference model. Sometimes including all the available features in an
elaborate model can be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018} can produce useful reference
models.

The remainder of the paper is structured as follows. In Section
\ref{reference-model-approach}, we review the concept of the reference
model, its benefits with examples and how it can be used as a filter
on data in a simple way. In Section \ref{comparison-minimal-subset} and 
Section \ref{comparison-complete-subset}, we show the
benefits of a reference model approaches for minimal and complete 
feature selection, respectively, before we illustrate the use 
of a reference model in a detailed case study of the 'body fat' data.
Finally, we end with a conclusion in Section \ref{conclusion}.

@@

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability and tends to
obscure the relevance of the covariates in relation to the target
variable of interest. We demonstrate with the following simple
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim \N(0,1) && \nonumber \\ 
     Y|&f\sim \N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} \N(\sqrt{\rho}f,1-\rho) \quad &&j=1,\ldots,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} \N(0,1) &&j=k+1,\ldots,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ covariates are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ covariates, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant covariate and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each covariate (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

When using a reference model, we first obtain predictions for $f$
using all the covariates $\{X_{j}\}_{j=1}^{p}$ taking into account
that we have only observed the noisy representation $Y$ of $f$. If our
model is good, we are able to describe $f$ better than $Y$ itself can,
which improves the accuracy in the estimation of the relevance of the
covariates.  Figure \ref{fig:correlation} illustrates this process in
the form of a scatter plot of (absolute) correlations of the
covariates with $Y$ against the corresponding correlations with the
predictions of a reference model (in this case the posterior
predictive means of model \eqref{eq:ref_mod}; see Section
\ref{simulations}). Looking at the marginal distributions, we see
that using a reference model to filter out noise in the data, the two
groups of features (relevant and non-relevant) can be distinguished
much better than when the correlation is computed using the observed
noisy data directly.

@@

In the bodyfat example above, we desired at the same time good
predictive performance and selection of a smaller number of
features. When the goal is to select a minimal subset of features,
which have similar predictive performance as the full model, we call it
\emph{minimal subset feature selection}. This minimal subset might
exclude features which have some predictive information about the
target, but given the minimal subset these features are not able to
provide such additional information that would improve predictive
performance significantly. The usual reason for this is that the relevant features
which are not in the minimal subset are highly correlated with features already in
the minimal subset. We will return to a problem of finding all
the features with some predictive power in Section~\ref{complete-selection}.

@@

In Figure \ref{fig:rmse_vs_fdr}, the predictive performance is shown
for different values of $n$ and $\rho$ in terms of RMSE and the false discovery rate (FDR, the ratio of the number of non-relevant selected
features over the number of selected features) of the selected submodel,
averaged after 100 data simulations. We see that projpred has superior FDR and the
smallest RMSE. Using a reference model improves steplm significantly,
but the minimal subset feature selection stays unreliable. Stepwise
Bayesian linear regression model without a reference model is better
than steplm without a reference model, but the minimal subset feature
selection is unreliable.

When the variable selection is repeated with different simulated data
sets, there is some variability in selected variables. We
measure the stability of variable selection by computing the entropy
of observed distribution over different models. The smallest entropy
would be obtained if the approach always selects the same set of
variables, and the largest entropy would be observed if the approach
would always select different set of variables.
Figure \ref{fig:entropy} shows the entropy scores for the different compared
methods. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy cannot distinguish the interchangeability 
due to correlation from instability. Thus, such a measure should be 
considerd as a relative and not as an absolute measure of stability.
The use of a reference model
improves slightly the stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

\hypertarget{real-world-data}{%
\subsection{Body fat example: Part 2}\label{real-world-data}}

Here we repeat the selection of Section \ref{intro-example} via
stepwise backward regression. In this case, the overall number of
covariates (original plus noisy) is 100, as it was in the last part of
Section \ref{intro-example}. We compare results with and without using
a simple reference model approach outlined in \eqref{eq:ref_mod} with
steplm. Figure \ref{fig:bodyfat_step_refvsdata} shows the number of
irrelevant features included in the final model and the out-of-sample
root mean square error (RMSE). Results are based on 100 bootstrap
samples on the whole dataset, and the predictive performance is tested
on the observations excluded at each bootstrap sample. We observe that
the reference model reduces the number of irrelevant features included
in the final model. This leads to less overfitting and thus to
improved out-of-sample predictive performance in terms of RMSE. The
reference model approach applied to the stepwise backward regression
achieves outstanding improvements considering its simplicity, yet it
does not reach the goodness of the much more sophisticated projective
prediction approach (see results of Section \ref{intro-example}).
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/bodyfat_step_refvsdata.pdf}
  \caption{Stepwise backward selection with and without using a reference model. The x-axis denotes the number of selected irrelevant features on the left and the out-of-sample RMSE on the right-hand side based on 100 bootstrap samples.\\}
  \label{fig:bodyfat_step_refvsdata}
\end{figure}

@@

An alternative to minimal subset feature selection is \emph{complete
  feature selection} in which the goal is to find all relevant
features that have some predictive information about the target. In
complete feature selection, it is possible that there are theoretically
relevant features, but given finite noisy data we are not able infer their relevance.
The projection predictive approach was originally designed for the minimal subset
feature selection, but we test here a simple iterative variant for the complete
feature selection case. In addition, we analyse the benefits of the reference
model approach in combination with three other
methods which have been originally designed for complete feature selection.

As the criteria of selection performance, we evaluate the average false
discovery rate and the average
sensitivity (i.e., the ratio of the number of relevant selected
features over the total number of relevant features). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \citet{paper:stability}.

@@

We modify the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ is non-trivial, and we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.

@@

% Many Bayesian approaches for complete feature selection first fit a
% Bayesian regression model using some sparsifying prior on the
% regression coefficients, for example, as discussed in
% \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and
% \cite{johnstone2004needles}.  In our experiments, we include the
% Dirichlet-Laplace \cite[DL;][]{paper:dirichlet_laplace} prior and the
% regularised horseshoe (RHS) prior \citep{paper:rhs}, which in our experience
% scales better to higher number of covariates as compared to other
% horseshoe-like priors.
% \note{Aki: How this paragraph connects to the next paragraph?}

@@

%Within the set of complete selection procedures, we consider three
%Bayesian methods.  \cite{johnstone2004needles} use a spike-and-slab
%type prior given by a mixture of a delta distribution in zero and a
%heavy-tailed distribution \cite[see
%also][]{paper:spike_slab_mitchell}. When the latter is a Laplace
%distribution, they show an interesting thresholding property of the
%median estimator and, based on this observation, present a complete
%selection procedure using their prior. We refer to this method as
%\emph{EB.med}. Further, we use a method to control for the local false
%discovery rate (\emph{loc.fdr}) as described in \cite{paper:efron,
%  efron2012large}. Finally, we device a simple selection procedure
%based on posterior credible intervals using the regularised horseshoe
%prior in which we select covariates whose 90\% credible intervals
%excludes zero (\emph{ci.90}) \note{Reference for this?}.

@@

The selection
by 90\% posterior credible intervals is done using the regularised
horseshoe prior and selecting those features whose posterior
distribution does not include zero in the interval between the 5\% and
the 95\% quantiles (cite).

All of these methods provide a complete selection approach and we
compare their performance with and without using a reference model.
That is, in the data condition, we apply the method on the original
data $y$ while, in the referece model condition, we replace $y$ by
their mean predictions $\hat{y}$ based on the reference model.

@@

Figure \ref{fig:sensitivity_vs_fdr_iterated} shows the average sensitivity on
the vertical axis and the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance, as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger data set size ($n$) and the higher the true correlations ($\rho$), the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant covariates. The iterative
projpred has good false discovery rate in all cases, and the
sensitivity is good except when the number of observations and the
correlation level are small. Tuning $\alpha$ might lead to improved
selection, yet the main source of poor performance could well be the
miscalibration of the uncertainty of the predictive utility score for
misspecified models with cross-validation \citep{bengio2004no}.

>>>>

    \author{
    Federico Pavone \footnote{Department of Computer Science, Aalto University, Finland.} $\,$\footnote{Correspondence concerning this article should be addressed to TODO.}, 
  Juho Piironen$^*$,
  Paul-Christian B\"{u}rkner$^*$
  and Aki Vehtari$^*$
  }

@@

\begin{document}
\maketitle
\begin{abstract}
  Feature selection, or more generally, model reduction is an important aspect of the statistical workflow aiming to provide insights from data. In this paper we discuss and demonstrate the benefits of using a reference model in feature selection. A reference model acts as a noise-filter on the target variable by modeling its data generating mechanism. As a result, using the reference model predictions in the model selection procedure reduces the variability and improves stability leading to improved model selection performance. Assuming that a Bayesian reference model describes the true distribution of future data well, the theoretically preferred usage of the reference model is to project its predictive distribution to a reduced model leading to projection predictive feature selection approach. Alternatively, reference models may also be used in an ad-hoc manner in combination with common feature selection methods.  In several numerical experiments, we investigate the performance of the projective prediction approach as well as alternative feature selection methods with and without reference models. Our results indicate that the use of reference models generally translates into better and more stable feature selection. Additionally, we demonstrate that the projection predictive approach shows superior performance as compared to alternative feature selection methods independently of whether or not they use reference models.
% These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.
\end{abstract}

@@

In statistical applications, one of the main steps in the modelling
workflow is feature selection, which is a special case of
model reduction. Feature selection (also known as variable or 
covariate selection) may have multiple goals.  First,
if the features themselves are of interest, we can use feature
selection to infer which features contain predictive information
about the target variable.  Second, as simpler models come with the
advantages of reduced measurement costs and improved interpretability,
we may be interested in finding the minimal subset of features which
still provides good predictive performance (or good balance between
simplicity and predictive performance).  When the predictive
capability is guiding the selection, the true data generation
mechanism of future data can be approximated either by using the
observed data directly or alternatively by using predictions from a
\emph{reference model} \citep{vehtari2012survey}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all features.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects our beliefs about
the future data in the best possible way and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian,gabry2019visualization}.  The reference model approach
has been used in Bayesian statistics at least since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see the papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see a paper by \citet{paper:projpred}.

Reference models have been also used in non-Bayesian contexts, where
\cite{harrell2015regression} describes them as full models that can be
thought of as a ``gold standard'' (for a given application).  For
example, \cite{faraggi2001understanding} deal with the necessity of
identifying interpretable risk groups in the context of survival data
using neural networks, which typically perform very well in terms of
prediction, but whose features are difficult to be understood in
terms of relevance.  \cite{paul2008preconditioning}, using the term
preconditioning, explore approximating models fitting Lasso or
stepwise regression against consistent estimates $\hat{y}$ of a
reference model instead of the observed responses $y$.
Whatever the terminology or applied statistical framework, reference
models offer a powerful approach to improving feature selection as we
will demonstrate in the present paper.

This paper is structured as follows. In Section
\ref{ref-intro}, we review the concept of the reference
model, its benefits with examples and how it can be used as a filter
on data in a simple way. In Section \ref{comparison-minimal-subset} and 
Section \ref{comparison-complete-subset}, we show the
benefits of a reference model approaches for minimal and complete 
feature selection, respectively, before we end with a conclusion in Section \ref{conclusion}.

\section{Reference models in feature selection}
\label{ref-intro}

@@

\subsection{Bodyfat example: Part 1}
\label{bodyfat-1}

To motivate the further discussion and experiments, we start by simple
feature selection example using body fat data by \citet{johnson1996fitting}.
We compare the projective prediction approach (\emph{projpred},
\citet{paper:projpred}) which uses a reference model, and classic
stepwise backward regression (steplm).
% \footnote{The analysis is based on \texttt{R}-notebook by Vehtari available at
%   \url{https://avehtari.github.io/modelselection/bodyfat.html}.}.
The experiments are implemented in \texttt{R} \citep{Rcore2018}.

The target variable of interest is the amount of body fat, which is
measured by immersing a person in water tank which is a complex and
expensive procedure. Additionally, we have information about 13
features which are anthropometric measurements (e.g., height, weight
and circumference of different body parts). The
features are highly correlated which causes additional
challenge in the feature selection. In total, we have 251
observations. The goal is to find the model which is able to predict
the amount of body fat well while requiring least amount of
measurements for a new person.

\cite{paper:bodyfat} report results using \emph{steplm} with a significance
level of 0.157 with AIC selection \citep{akaike1974new}, fixing
abdomen and height to be always included in the model. For better
comparison we don't fix any of the variables.  The steplm approach is
carried out combining the \texttt{step} and \texttt{lm} functions in R.

For the selection via projpred, the Bayesian reference model includes
all the features using a regularised horseshoe prior~\citep{paper:rhs} on the
feature coefficients. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using approximate leave-one-out cross-validation via Pareto-smoothed importance-sampling
\citep[PSIS-LOO-CV; ][]{paper:psis_loo}.  We select the smallest  submodel with an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach, and papers by \citet{paper:model_selection} and
\citet{paper:projpred} for more details. The complete projpred approach is implemented in the \texttt{projpred} R package.

@@

We repeat the experiment with a modified data set by adding 84
unrelated noisy features so that there are 100 feature in total. 
The last two rows of Table
\ref{tab:model_performances} show the cross-validated RMSE, the size
of the selected model and the number of selected noisy features using
projpred or steplm. The results show that projpred has similar
predictive performance and the same number of selected features as
with the original data, whereas the stepwise regression has worse
predictive performance and the number of selected features is much
higher and include a large number of irrelevant features.

@@

Both projpred and steplm compare a large number of models using either
forward or backward search, which can lead to selection induced
overfitting, but even with 100 features, projpred is able to select a
submodel with similar performance as the full model. In this example, the two
 compared methods also differ in other aspects than the usage of a reference 
 model, such as that projpred
uses Bayesian inference and steplm uses maximum likelihood estimation. 
However, as we show in the present paper, one of the primary difference 
with respect to quality of the feature selection is indeed whether or 
not a reference model is applied.

@@

If one of the models to be compared is the full model which can be
used as a reference model, there is no additional cost of using a
reference model as it was estimated as part of the analysis anyway.
Sometimes including all the available features in an
elaborate model can be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018} can produce useful reference
models.

@@

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability and tends to
obscure the relevance of the features in relation to the target
variable of interest. We demonstrate with the following simple
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim \N(0,1) && \nonumber \\ 
     Y|&f\sim \N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} \N(\sqrt{\rho}f,1-\rho) \quad &&j=1,\ldots,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} \N(0,1) &&j=k+1,\ldots,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ features are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ features, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant feature and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each feature (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

When using a reference model, we first obtain predictions for $f$
using all the features $\{X_{j}\}_{j=1}^{p}$ taking into account
that we have only observed the noisy representation $Y$ of $f$. If our
model is good, we are able to describe $f$ better than $Y$ itself can,
which improves the accuracy in the estimation of the relevance of the
features.  Figure \ref{fig:correlation} illustrates this process in
the form of a scatter plot of (absolute) correlations of the
features with $Y$ against the corresponding correlations with the
predictions of a reference model (in this case the posterior
predictive means of model \eqref{eq:ref_mod}; see Section
\ref{simulations}). Looking at the marginal distributions, we see
that using a reference model to filter out noise in the data, the two
groups of features (relevant and non-relevant) can be distinguished
much better than when the correlation is computed using the observed
noisy data directly.

@@

In the bodyfat example above, our two simultaneous goals were to 
obtain good predictive performance and to select a smaller number of features. 
When the goal is to select a minimal subset of features,
which have similar predictive performance as the full model, we call it
\emph{minimal subset feature selection}. This minimal subset might
exclude features which have some predictive information about the
target but, given the minimal subset, these features are not able to
provide such additional information that would improve predictive
performance in a substantial manner.
The usual reason for this is that the relevant features
which are not in the minimal subset are highly correlated with features already in
the minimal subset. We will return to a problem of finding all
the features with some predictive power in Section~\ref{complete-selection}.

@@

In Figure \ref{fig:rmse_vs_fdr}, the predictive performance is shown
for different values of $n$ and $\rho$ in terms of RMSE and the false discovery rate (FDR, the ratio of the number of non-relevant selected
features over the number of selected features) of the selected submodel,
averaged after 100 data simulations. We see that projpred has superior FDR and the
smallest RMSE. Using a reference model improves steplm significantly,
but the minimal subset feature selection stays unreliable. Stepwise
Bayesian linear regression is better than steplm when comparing both methods with
and without a reference model, respectively. However, the minimal subset feature
selection of the Bayesian linear regression is still less reliable than projpred.

When the feature selection is repeated with different simulated data
sets, there is some variability in the selected features. We
measure the stability of feature selection by computing the entropy
of observed distribution over different models. The smallest entropy
would be obtained if the approach always selected the same set of
features, and the largest entropy would be observed if the approach
would always selected different sets of features.
Figure \ref{fig:entropy} shows the entropy scores for the different compared
methods. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
selected alternately, thus making stability estimation of the
selection a non-trivial task. Entropy cannot distinguish the interchangeability 
due to correlation from instability. Thus, such a measure should be 
considerd as a relative and not as an absolute measure of stability.
The use of a reference model
improves the stability of steplm in feature selection slightly, while it makes
little difference for the Bayesian linear regression. The
projpred approach turns out to be far more stable than all
other methods. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

\hypertarget{bodyfat-2}{%
\subsection{Body fat example: Part 2}\label{bodyfat-2}}

Here we repeat the selection of Section \ref{bodyfat-1} via
stepwise backward regression. In this case, the overall number of
features (original plus noisy) is 100, as it was in the last part of
Section \ref{bodyfat-1}. We compare results with and without using
a simple reference model approach outlined in \eqref{eq:ref_mod} with
steplm. Figure \ref{fig:bodyfat_step_refvsdata} shows the number of
irrelevant features included in the final model and the out-of-sample
root mean square error (RMSE). Results are based on 100 bootstrap
samples on the whole dataset, and the predictive performance is tested
on the observations excluded at each bootstrap sample. We observe that
the reference model reduces the number of irrelevant features included
in the final model. This leads to less overfitting and thus to
improved out-of-sample predictive performance in terms of RMSE. The
reference model approach applied to the stepwise backward regression
achieves outstanding improvements considering its simplicity, yet it
does not reach the goodness of the much more sophisticated projective
prediction approach (see results of Section \ref{bodyfat-1}).
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/bodyfat_step_refvsdata.pdf}
  \caption{Stepwise backward selection with and without using a reference model. The x-axis denotes the number of selected irrelevant features on the left and the out-of-sample RMSE on the right-hand side based on 100 bootstrap samples.\\}
  \label{fig:bodyfat_step_refvsdata}
\end{figure}

@@

An alternative to minimal subset feature selection is \emph{complete
  feature selection} in which the goal is to find all relevant
features that have some predictive information about the target. In
complete feature selection, it is possible that there are theoretically
relevant features, but given finite noisy data we are not able infer their relevance.
The projection predictive approach was originally designed for the minimal subset
feature selection but we will test a simple iterative variant for the complete
feature selection case in the section.
 In addition, we analyse the benefits of the reference
model approach in combination with three other
methods which have been specifically designed for complete feature selection. 
As the criteria of selection performance, we evaluate the average false
discovery rate and the average
sensitivity (i.e., the ratio of the number of relevant selected
features over the total number of relevant features). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \citet{paper:stability}.

@@

We modify the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model's predictive performance. 
 The iterative projection repeats
the projpred selection for different iterations, at each time
excluding the features selected in the previous
iterations from the search.
 At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stands for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ is non-trivial, and we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package, that is, $\alpha=0.16$.

@@

% Many Bayesian approaches for complete feature selection first fit a
% Bayesian regression model using some sparsifying prior on the
% regression coefficients, for example, as discussed in
% \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and
% \cite{johnstone2004needles}.  In our experiments, we include the
% Dirichlet-Laplace \cite[DL;][]{paper:dirichlet_laplace} prior and the
% regularised horseshoe (RHS) prior \citep{paper:rhs}, which in our experience
% scales better to higher number of features as compared to other
% horseshoe-like priors.
% \note{Aki: How this paragraph connects to the next paragraph?}

@@

%Within the set of complete selection procedures, we consider three
%Bayesian methods.  \cite{johnstone2004needles} use a spike-and-slab
%type prior given by a mixture of a delta distribution in zero and a
%heavy-tailed distribution \cite[see
%also][]{paper:spike_slab_mitchell}. When the latter is a Laplace
%distribution, they show an interesting thresholding property of the
%median estimator and, based on this observation, present a complete
%selection procedure using their prior. We refer to this method as
%\emph{EB.med}. Further, we use a method to control for the local false
%discovery rate (\emph{loc.fdr}) as described in \cite{paper:efron,
%  efron2012large}. Finally, we device a simple selection procedure
%based on posterior credible intervals using the regularised horseshoe
%prior in which we select features whose 90\% credible intervals
%excludes zero (\emph{ci.90}) \note{Reference for this?}.

@@

The selection
by 90\% posterior credible intervals is done using the regularised
horseshoe prior \Citep{paper:rhs} and selecting those features whose posterior
distribution does not include zero in the interval between the 5\% and
the 95\% quantiles.

All of these methods provide a complete selection approach and we
compare their performance with and without using a reference model.
That is, in the data condition, we apply the method on the original
data $y$ while, in the reference model condition, we replace $y$ by
their mean predictions $\hat{y}$ based on the reference model.

@@

Figure \ref{fig:sensitivity_vs_fdr_iterated} shows the average sensitivity on
the vertical axis and the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance, as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger data set size ($n$) and the higher the true correlations ($\rho$), the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant features. The iterative
projpred has good false discovery rate in all cases, and the
sensitivity is good except when the number of observations and the
correlation level are small. Tuning $\alpha$ might lead to improved
selection, yet the main source of poor performance could well be the
miscalibration of the uncertainty of the predictive utility score for
misspecified models with cross-validation \citep{bengio2004no}.
